---
title: "Access AIMS eReefs data with R"
output:
  html_notebook: default
  html_document:
    df_print: paged
---

This tutorial is based on the article [How to open and work with NetCDF data in R](https://rpubs.com/boyerag/297592).

## Load the required R packages
First of all, load the necessary libraries.

```{r results="hide"}
library(ncdf4) # package for netcdf manipulation
library(raster) # package for raster manipulation
library(ggplot2) # package for plotting
library(rgdal) # package for geospatial analysis
```
## Define which data to be plotted.
In this section we define which data we want to read and plot.

- **inputFile**  
  The netCDF input file. This can either be a downloaded file (see [How to manually download derived data from THREDDS](http://ereefs.aims.local/ereefs-aims/help/how-to-manually-download-derived-data)) or a  OPeNDAP URL from the [AIMS THREDDS server](http://thredds.ereefs.aims.gov.au/thredds/catalog.html). For this tutorial we are using the OPeNDAP URL for the file "gbr4_v2-daily-monthly-2010-09.nc".
- **selectedVariable**  
  The name of the variable in the netCDF file.
- **selectedTimeIndex**  
  The time slice in the netCDF file. For example, in the netCDF file "gbr4_v2-daily-monthly-2010-09.nc" the time steps are "days". This means if you select `selectedTimeIndex=2` it refers to the second day in this file, which is 02/09/2010.
- **selectedDepthIndex**  
  The depth slice in the netCDF file. See the following table for a mapping of index to value:

  Index (k) | Hydrodynamic 1km model | Hydrodynamic and BioGeoChemical 4km model
  ---|---|---
  1 | -140.00 m | -145.00 m
  2 | -120.00 m | -120.00 m
  3 | -103.00 m | -103.00 m
  4 | -88.00 m | -88.00 m
  5 | -73.00 m | -73.00 m
  6 | -60.00 m | -60.00 m
  7 | -49.00 m | -49.00 m
  8 | -39.50 m |-39.50 m
  9 | -31.00 m | -31.00 m
  10 | -24.00 m | -23.75 m
  11 | -18.00 m | -17.75 m
  12 | -13.00 m | -12.75 m
  13 | -9.00 m | -8.80 m
  14 | -5.25 m | -5.55 m
  15 | -2.35 m | -3.00 m
  16 | -0.50 m | -1.50 m
  17 | n/a | -0.50 m

```{r}
# OPeNDAP URL to file EREEFS_AIMS-CSIRO_gbr4_v2_hydro_daily-monthly-2010-09.nc. Hydrodynamic 4km model, daily data for the month September 2010
input_file <- "http://thredds.ereefs.aims.gov.au/thredds/dodsC/s3://aims-ereefs-public-prod/derived/ncaggregate/ereefs/gbr4_v2/daily-monthly/EREEFS_AIMS-CSIRO_gbr4_v2_hydro_daily-monthly-2010-09.nc"
input_file

# The "temperature" variable
selectedVariable <- "temp"
selectedVariable

# 2nd of September 2010
selectedTimeIndex <- 2
selectedTimeIndex

# -1.50 m depth
selectedDepthIndex <- 16
selectedDepthIndex
```
## Read in the netCDF file contents
Read in the netCDF file contents and store the latitude, longitude, time and depth data in variables for later use.
```{r}
nc_data <- nc_open(input_file)

lon <- ncvar_get(nc_data, "longitude")
head(lon) # look at the first few entries in the longitude vector

lat <- ncvar_get(nc_data, "latitude")
head(lat) # look at the first few entries in the latitude vector

time <- ncvar_get(nc_data, "time")
head(time) # look at the first few entries in the time vector

depth <- ncvar_get(nc_data, "zc")
head(depth) # look at the first few entries in the depth vector
```
Read in the data from the selected variable and select the time and depth slice.
```{r}
variableData.slice <- ncvar_get(nc = nc_data,
                                varid = selectedVariable,
                                start = c(1, 1, selectedDepthIndex, selectedTimeIndex),
                                count = c(length(lon), length(lat), 1, 1))
dim(variableData.slice)

```
Now that all data is stored in memory, close the netCDF file.
```{r}
nc_close(nc_data)
print("file closed")
```
## Working with the data

Save the data in a raster. We also need to transpose and flip to orient the data correctly.
```{r}
r <- raster(t(variableData.slice), xmn = min(lon), xmx = max(lon), ymn = min(lat), ymx = max(lat), crs = CRS("+init=epsg:4326"))
r <- flip(r, direction = 'y')
dim(r)
```
## Plotting
Finally plot the data.

```{r}
plot(r)
```
